---
title: "rockfishr"
subtitle: "Common code for ABL rockfish assessments"
author: Benjamin Williams (ben.williams@noaa.gov)
date:    
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rockfishr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Goal

The goal of rockfishr is to create a clear workflow for pulling and cleaning data for rockfish fishery stock assessments based upon the ADMB rockfish assessment model. 
It utilizes a "project oriented workflow" via RStudio. 
You must be able to have a connection to the AFSC & AKFIN (Answers) data servers (e.g., VPN if offsite), and have usernames/passwords setup.

## Installation
The `rockfishr` package can be installed from [github](https://github.com/BenWilliams-NOAA/rockfishr) with:
```{r}
# install.packages("devtools")
devtools::install_github("BenWilliams/rockfishr")
```


## Use

There are a suite of "global" items that are used repeatedly in the `rockfishr` package. 
Currently the package is setup to deal with northern rockfish "NORK" or dusky rockfish "DUSK", other species can be added as needed. 


```{r setup}
library(rockfishr)

# globals ----

species = "NORK" # northern rockfish
year = 2020 # assessment year
afsc_user = "your_afsc_username"
afsc_pwd = "your_afsc_password"
akfin_user = "your_akfn_username"
akfin_pwd = "your_akfin_password"

admb_home = "C:/Program Files (x86)/ADMB-12.1" # location admb exists on your computer 
                                               # if is "c:/admb" can leave NULL
TAC = c(3786, 3681, 4528) # last three years of TAC (year-3, year-2, year-1)
rec_age = 2 # recruitment age
plus_age = 45 # plus group age
model_name = "updated_nr" # name of ADMB .tpl file
dat_name = "goa_nr_2020" # name of ADMB .dat file

mcmc = 1e+07 # number of mcmc draws
mcsave = 2000 # number of saved mcmc draws

```

### Create directories

The first function creates a number of directories within your R project folder. The setup is organized by year within a folder. 
Using the globals above in the `modeldir()` function would create the following folders and files:

```{r}
# setup folders -----

modeldir(year)

```
```
project
|__2020
    |__data
        |__models
        |    |__ageage
        |    |     AGEAGE.tpl
        |    |__allometric
        |    |     allometric.tpl
        |    |__length_sd
        |    |     lengthSD.tpl
        |    |__VBL
        |    |      VBL.tpl
        |    |__wVBL
        |    |      wVBL.tpl
        |    |      lvb.ctl
        |
        |__output
        |__raw
        |__SARA
        |__user_input

```

This can then be updated for each year. 

```{r}
# setup folders -----

modeldir("2021")

```

Which would create a duplication of the folders for the given year

```
project
|__2020
|   |__data
|       |__models
|       |    |__ageage
|       |    |     AGEAGE.tpl
|       |    |__allometric
|       |    |     allometric.tpl
|       |    |__length_sd
|       |    |     lengthSD.tpl
|       |    |__VBL
|       |    |      VBL.tpl
|       |    |__wVBL
|       |    |      wVBL.tpl
|       |    |      lvb.ctl
|       |
|       |__output
|       |__raw
|       |__SARA
|       |__user_input
|__2021
    |__data
        |__models
        |    |__ageage
        |    |     AGEAGE.tpl
        |    |__allometric
        |    |     allometric.tpl
        |    |__length_sd
        |    |     lengthSD.tpl
        |    |__VBL
        |    |      VBL.tpl
        |    |__wVBL
        |    |      wVBL.tpl
        |    |      lvb.ctl
        |
        |__output
        |__raw
        |__SARA
        |__user_input        

```

### Query raw data

The next step is to query the raw data from the AKFIN and AFSC databases. 

```{r}
# query databases ----
raw_data(species, year, afsc_user, afsc_pwd, akfin_user, akfin_pwd)
```

The initial output will be raw data pulls that will be placed in the `year/data/raw` folder.
These are kept in the "raw" format with an associated timestamp of when the data were queried.
To see the queries being perfomed one can either type `raw_data` in their console and hit `enter`, or explore the code on [github](https://github.com/BenWilliams-NOAA/rockfishr/blob/main/R/raw_data.R).


```
project
|__2020
    |__data
        |__models
        |__output
        |__raw
            data_called.txt
            fishery_age_comp_data.csv
            fishery_catch_data.csv
            fishery_obs_data.csv
            fishery_size_comp_freq.csv
            srv_age_comp.csv
            srv_age_specimens.csv
            srv_biomass.csv
            srv_saa_age.csv
            srv_saa_length.csv
            srv_size_comp.csv
            srv_size_freq.csv
```

### Clean and process data

There are a suite of functions to process the raw data, the results are placed in the `output` folder.
These functions need to be run in order.

```{r}
clean_catch(year, TAC)
ageage(reader_tester = NULL, species, year, admb_home, region = "GOA", rec_age, plus_age)
fish_age_comp(year, rec_age, plus_age)
survey_age_comp(year, rec_age, plus_age)
fish_size_comp(year, rec_age)
survey_size_comp(year)
size_at_age(year, admb_home, rec_age)
weight_at_age(year, admb_home, rec_age)  
# note: must provide a file for VAST survey biomass estimates otherwise D-B estimates are output
survey_biomass(year) 
concat_dat(year, "db", rec_age, plus_age) # create a .dat file in the "db" folder

# VAST "bridge model"
survey_biomass(year, "VAST_estimate_mesa.csv") 
concat_dat(year, "m18.2", rec_age, plus_age) # create a .dat file in the "m18.2" folder
```  

Alternatively there is a function that will run all of these functions

```{r}
clean_raw(year, species, TAC, admb_home, rec_age, plus_age, reader_tester = NULL)
```


### Run models

Currently all models are being run through the ADMB command line.  
Compile model, run model with -mcmc and -mcsave (if running mcmc...).  
Run model with -mceval.  

### Process results

Define the model (e.g., folder) run the `process_results` function and the `base_plots` function. 

The `process_results` function cleans outputs from ADMB and generates a series of .csv files that are placed within a folder titled `processed` nested in the model folder. 
The function also creates a `figs` and `tables` folder within the model folder.

The plotting function will populate the `figs` folder with a number of figures from the model output. 

```{r}
model = "db"
process_results(year, model, model_name, data_name, rec_age, plus_age, 
                mcmc = mcmc, mcsave = mcsave)

base_plots(year, model, model_name, rec_age)


model = "m18.2"

process_results(year, model, model_name, data_name, rec_age, plus_age, 
                mcmc = mcmc, mcsave = mcsave, survey = "VAST_estimate_mesa.csv")

base_plots(year, model, model_name, rec_age)
```


```
project
|__2020
    |__data
    |__db
    |   |__figs
    |   |__processed
    |   |__tables
    |__m18.2
        |__figs
        |__processed
        |__tables        
```

## Retrospective

For any models that you have run mcmc draws on and would like retrospective examinations the following code can be used to run the retros. 
The output will be placed in a `retro` folder in the model directory.

```{r}
# run retro ----
run_retro(year, model = "m18.2", tpl_name = model_name, n_retro = 10,
          admb_home = admb_home, mcmc = 500000, mcsave = 100)
```

Retrospective results can then be plotted. 
These figures will be located in the `figs` folder.

```{r}
# plot retros ----
plot_retro(year, model)
plot_retro_survey(year, model)
```

